### 2015-10-01 manuscript.Rnw
add example on how to use rotl with rgbif

>>>>

In recent years, R has become part of the toolbox of many researchers in
evolutionary biology. As many methods for comparative analyses are implemented,
and because it is a relatively easy to use programming language, R greatly
facilitates the analysis of large datasets, and allows researchers to combine
methods in novel ways. Additionally, as more data is made available online and
accessible using APIs, many R packages have been developed to take advantage of
these datasets, especially spearheaded by the organization rOpenSci
\url{https://ropensci.org}. \texttt{rotl} contributes to these initiatives and
data from OTL can be combined with other sources easily. For instance in
Figure~\ref{cat_map}, we show how we can obtain a map of the occurrences for
some of the cat species (genus \textit{Lynx}) included in the phylogeny above
using records for these species found in GBIF (code is available in
supplementary material).

<<cat_map, echo=FALSE, results='hide', message=FALSE, cache=TRUE, fig.cap='GBIF records for the species in \\textit{Lynx} included in the phylogeny associated with the study by \\citealt{Johnson2006}'>>=

suppressPackageStartupMessages(library(rgbif))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(maps))

cat_species <- cat_tree$tip.label
lynx_species <- grep("^Lynx", cat_tree$tip.label, value = TRUE)

gbif_keys <- sapply(lynx_species,
                    function(x) name_backbone(name = x)$speciesKey,
                    USE.NAMES = FALSE)

lynx_loc <- occ_search(taxonKey = gbif_keys, limit = 500,
                       return = "data", fields = "minimal",
                       hasCoordinate = TRUE)

lynx_loc <- do.call("rbind", lynx_loc)
lynx_loc[["decimalLatitude"]] <- as.numeric(lynx_loc[["decimalLatitude"]])
lynx_loc[["decimalLongitude"]] <- as.numeric(lynx_loc[["decimalLongitude"]])
lynx_loc <- lynx_loc[complete.cases(lynx_loc), ]

names(lynx_loc)[1] <- "Species"

world <- map_data("world")

ggplot(lynx_loc) +
    annotation_map(world, fill="gray40", color="gray40") +
    geom_point(aes(y = decimalLatitude, x = decimalLongitude, color = Species),
               alpha = 0.3, size = 2) +
    coord_map(projection = "mercator", orientation = c(90, 0, 0)) +
    xlab("Longitude") + ylab("Latitude")

@
